The conclusion of this master thesis marks the end of an amazing and very intense period of my life. My adventure in Norway has come to an end and a new beginning is opening in front of me. I feel like I ought to thank all the great people I have met and worked with in these three years. 

First I have to thank my two supervisors: Morten Hjorth-Jensen and Andrea Shindler. I want to express my deep admiration for the constant support and the endless inputs even if we were on two different continents. I really thank you for the opportunity you have given me and for the time spent at MSU together.

Then I want to thank Mathias Vege, for having shared this journey into Lattice QCD with me. I don't know what I would have done without all the useful discussions we had. I really enjoyed our adventure in Michigan and I hope we'll work again together in the future.

Next, I have to thank Alessio Pizzini. You know how much it meant to have someone to share this three years exploring Physics and Norway. 

I also have to thank Anders Johansson, my office mate. It's been great sharing with you the most random thoughts every day, together with the weirdest pieces of code we could find. Everything in between a game of Smash and the other.

A big thank you to all the Comp-Phys group at UiO. Spending my time with you has been amazing, no matter if it was on Physics, playing in the kitchen or talking about the most diverse issues. I loved spending my week-ends, evenings and nights at work because of you. I hope the spirit of this group will never be lost.

I cannot disentangle this work from all the other people that made my stay in Norway so nice. Thanks to Davide, Lorenzo and Federico, but also Giulio and Mattia. You guys have been amazing. Thanks also to Jan, Diane and Marie, because our friendship has been something special for me. And thanks to all the other great people I have met here.

Then to my parents, because they filled me with curiosity, inspired me and made me who I am. For the unconditional support you've given me. Regardless of the distance you always help me.

And finally to you two, because no matter where life is taking us, you are the people I care about the most.

\vspace{2cm}
\begin{flushright}
    Oslo, 14th May 2018\\
    Giovanni Pederiva    
\end{flushright}




