
Lattice QCD is one of the main Lattice Field Theories. It deals with the strong force at low energies, handling non-perturbative calculations in a numerical way. The idea of discretizing spacetime in a lattice and perform calculations of field theories was proposed by Wilson in 1974 \cite{wilson_confinement_1974} as an alternative method to study the confinement of quarks. It has proven to be the most systematic approach to non-perturbative theories like QCD. In this chapter we will describe briefly how this discretization procedure is performed and some of the main computational strategies that are involved.

\section{Discretizing Field Theories}

The starting point for Lattice QCD is Feynman's path-integral formalism, but expressed in Euclidean spacetime, through a Wick rotation. An observable of some field $\phi$ is then given by:
\beq
\braket{O[\phi]} = \frac{1}{Z[\phi]} \int \D[\phi] ~O[\phi] e^{-S_E[\phi]},
\label{lattice:obs_base}
\eeq
and the euclidean correlator of two quantities $O_1[\phi]$ and $O_2[\phi]$ between times $t_1$ and $t_2$ is:
\beq
\braket{O_1[\phi](t_1)O_2[\phi](t_2)} = \frac{1}{Z[\phi]} \int \D[\phi] ~O_1[\phi(t_1)]O_2[\phi(t_2)] e^{-S[\phi]},
\label{lattice:correlator}
\eeq
where the partition function $Z[\phi]$ is defined as:
\beq
	Z = \int \D[\phi] ~e^{-S_E[\phi]} ,
\eeq
and $S_E[\phi]$ is the euclidean action of the field:
\beq
S_E[\phi] = \int d^4x \Lagr [\phi] ,
\eeq
Evaluating path-integrals is not possible in general with analytical tools so, in order to allow numerical computations, the Euclidean spacetime is discretized on a lattice, fixing the value of the field at only specific and equally spaced position in spacetime. 


\subsection{The Quantum Mechanical Example}
To better understand the discretization procedure for quantum field theories, it is useful to look at the simpler case of a quantum mechanical system, for example of a single particle in a potential $V$. The Hamiltonian of the system is simply:
\beq
    \hat H = \frac{\hat p^2}{2m} + \hat V,
\eeq
where we used natural units ($\hbar = c =1$). In path integral formalism the propagator between a state $\ket{x_i}$ at $t_i$ to another state $\ket{x_f}$ at $t_f$ can be expressed as:
\beq
    \bra{x_f} e^{-\hat H(t_f-t_i)} \ket{x_i} = \int \D x(t)~e^{-S_E[x]},
    \label{eq:action__}
\eeq
with $S_E[x]$ being the euclidean action:
\beq
    S_E[x(t)] = \int_{t_i}^{t_f} dt \Lagr [x(t)] = \int_{t_i}^{t_f} dt \left(\frac{m}{2}\dot x(t)^2 +  V[x(t)]\right).
\eeq
Note that the in the path integral in \cref{eq:action__} the exponent contains no imaginary unit. This is because we are using the euclidean action, related to the classical action by a Wick rotation $it \rightarrow t$. \\
To compute the propagator numerically one can chose to discretize the time dimension into slices $t_0, t_1,\dots,t_N$ of distance $a$ one from another. A path then becomes a vector of positions at fixed time points:
\beq
    x = \{x(t_0),x(t_1), \dots, x(t_N)\}.
\eeq
The integral is then changed from a continuous definition to an $N-2$ dimensional integral (the boundary points are fixed):
\beq
\int \D x(t)~e^{-S_E[x](t)} \rightarrow \frac{1}{A}\int dx_1 dx_2\dots dx_{N-1}~e^{-S_{lat}[x]} ,
\eeq
where $A$ is some normalization. The classical action is now only computed on the fixed points of the discretized time coordinate. In particular we can approximate the integral between $t_j$ and  $t_{j + 1}$, using the finite difference discretization for the derivative of $x(t)$ and the average potential, as:
\beq
    \int_{t_j}^{t_{j+1}} dt \Lagr[x(t)] \approx a\left[ \frac{m}{2} \left(\frac{x_{j+1} - x_j}{a} \right)^2 + \frac{1}{2} (V[x_{j_+1}] - V[x_j]) \right].
\eeq
With this approximation the action of one path, or ``configuration'', is the sum of the approximated integrals from:
\beq
S_{lat}[x] \approx \sum_{i = 0}^{N-1} \left[ \frac{m}{2a} (x_{j+1} - x_j)^2 + aV(x_i)\right].
\eeq
This discretization procedure for a simple quantum mechanical system (whose space coordinate can be seen as a one-dimensional scalar field) allows us to compute the path integrals in \cref{lattice:obs_base,lattice:correlator} using, in the case for the expectation value of $O[x]$:
\begin{align}
    \langle O[x] \rangle &=  \frac{1}{Z[x]} \int \D[x] ~O[x] e^{-S_{E}[x]} = \frac{1}{Z[x]} \int dx_1 dx_2\dots dx_{N-1}~O[x]e^{-S_{lat}[x]}  \\\nonumber
    &=\int dx_1 dx_2\dots dx_{N-1} P[x] O[x],
\end{align}
where we have identified $P[x] = \frac{1}{Z[x]}e^{-S_{lat}[x]}$ as the probability distribution function of the observable $O$. One can then compute the expectation value numerically by choosing random configurations according to the specified PDF, that depends on the action and compute the expectation value as the average of the observable over the chosen configurations.


\section{Discretization of QCD on the Lattice}
\label{sec:lattice_discretize}
In the case of QCD the discretization procedure is more complicated than the one shown in the previous section for a one dimensional harmonic oscillator, chiefly because there are two types of fields at play. Moreover the gluon gauge field $A$ and the $N_f$ fermionic quark fields $\psi$ are defined in four-dimensional spacetime. In order to perform numerical calculations on the lattice, it is necessary to discretize the Euclidean action for both fields on a hyper-cubic lattice $\Lambda$:
\beq
    \Lambda = \{n = (n_x, n_y, n_z, n_t)\}~~\text{where}~~n_x, n_y, n_z = 1,2,\dots,N;~~n_t = 1,2,\dots,N_T.
\eeq
The choice of the lattice spacing, usually denoted $a$, is arbitrary, but most often it is chosen to be equal for all dimensions, space and time alike. Our discretized fields are then constrained to have values only on the points $an$ instead of on a continuum spacetime $x^\mu$:
\beq
    \phi(x) \xrightarrow{discretization} \phi(an).
\eeq
In this section the discretization is performed following a structure that resembles \cref{intro:symmetry}, that is to start from the fermion fields and imposing gauge invariance.

\subsection{Na{\"i}ve Discretization of Fermions}
\label{lat:ferm_naive}
The main ideas for the discretization procedure of fermions that we present are taken from \cite{gattringer_quantum_2010}, quoting the main intermediate steps that lead to the formulation of Lattice QCD. The reviews \cite{gupta_introduction_1998,lepage_lattice_2005} are useful references.\\
The starting point is the fermionic euclidean action in the continuum:
\beq
    S_F [\psi,\bar\psi] = \int d^4x \bpsi(x)(\gamma_\mu\partial^\mu + m)\psi(x).
\eeq 
We discretize the euclidean spacetime on a lattice $\Lambda$ of spacing $a$, each point will be denoted with $n$. The partial derivative can be turned into the central finite difference between neighboring points along the direction of the derivative. This means to take the difference between the value of the field $\psi(n)$ at its neighboring point along the direction $\hat\mu$ and dividing by the distance between them, two times the lattice spacing $a$:
\beq
    \partial_\mu\psi(x) \rightarrow \frac{\psi(n+\hat\mu) - \psi(n-\hat\mu)}{2a}.
\eeq 
The discretized fermion action is then the sum of the lagrangian density over all lattice sites multiplied by the unit volume $a^4$:
\beq
    S_F [\psi,\bar\psi] = a^4 \sum_{n\in\Lambda} \bpsi(n) \left[ \sum_{\mu=1}^4\gamma_\mu  \frac{\psi(n+\hat\mu) - \psi(n-\hat\mu)}{2a} + m \psi(n) \right].
\eeq 
As we did in \cref{intro:symmetry}, we try to apply a local gauge transformation $\Omega(n)$ to the field. It is simple to show that some terms of the derivative in the action are not gauge invariant:
\beq
    \bpsi(n)\psi(n\pm\hat\mu) \rightarrow \bpsi(n)\Omega^\dagger(n)\Omega(n\pm\hat\mu)\psi(n\pm\hat\mu).
    \label{eq:fermion_wrong}
\eeq
The problem is in the way the derivative was discretized and the solution is the introduction of an additional field that has the correct transformation laws. This field must transform under a local gauge transformation in a way that connects the values of the fermion field at two different lattice sites. Because of this property we denote the auxiliary field as ``link variables'', later the matching with the gauge field will be introduced, and by convention it is written as $U(n, n + \hat\mu) = U_\mu(n)$. Furthermore, it must depend on the orientation along the $\mu$ direction which from \cref{eq:fermion_wrong} reads:
\beq
U(n, n-\hat\mu) \equiv  U_{-\mu}(n) = U_\mu(n-\hat\mu)^\dagger .
\eeq
In particular, we want it to transform as:
\begin{align}
    \bpsi(n)U_\mu(n)\psi(n+\hat\mu) &\rightarrow \bpsi(n)\Omega^\dagger(n)U'_\mu(n)\Omega(n+\hat\mu)\psi(n\pm\hat\mu)\\\nonumber
    \bpsi(n)U_{-\mu}(n)\psi(n-\hat\mu) &\rightarrow \bpsi(n)\Omega^\dagger(n)U'_{-\mu}(n)\Omega(n+\hat\mu)\psi(n\pm\hat\mu),
\end{align}
from which we infer that:
\begin{align}
    U_\mu(n)    &\rightarrow U'_\mu(n) = \Omega(n)U_\mu(n)\Omega^\dagger(n+\hat\mu)\label{link_transformation}\\\nonumber
    U_{-\mu}(n) &\rightarrow U'_{-\mu}(n) = \Omega(n)U_{-\mu}(n)\Omega^\dagger(n-\hat\mu).
\end{align}
the field $U_\mu(n)$ is an $\mathrm{SU}(3)$ valued field, that can be intuitively thought of as a set of oriented links between the sites of the lattice. A visual representation is given in \cref{fig:links},  
\begin{figure}[!htb]
\centering
\begin{tikzpicture}[node distance=1cm] 
    % arrows
    \draw[-<-=.5, line width=1.5pt, >=latex] (-2.5,0) to (-6.5,0);
    \draw[-<-=.5, line width=1.5pt, >=latex] (2.5,0) to  (6.5,0);
    
    \draw (6.5, 0.3) -- (6.5, -0.3);
    \draw (2.5, 0.3) -- (2.5, -0.3);
    \draw (-2.5, 0.3) -- (-2.5, -0.3);
    \draw (-6.5, 0.3) -- (-6.5, -0.3);

    \draw (6.5, 0) -- (6.8, 0);
    \draw (2.5, 0) -- (2.2, 0);
    \draw (-2.5, 0) -- (-2.2, 0);
    \draw (-6.5, 0) -- (-6.8, 0);

    \filldraw (6.5, 0) circle (0.05);
    \filldraw (2.5, 0) circle (0.05);
    \filldraw (-2.5, 0) circle (0.05);
    \filldraw (-6.5, 0) circle (0.05); 

    \node [above right] (A) at (-2.5, 0) {$n+\hat\mu$};
    \node [above left] (B) at (-6.5, 0) {$n$};
    \node [above right] (D) at (6.5, 0) {$n$};
    \node [above left] (C) at (2.5, 0) {$n-\hat\mu$};

    \node (E) at (4.5, -0.7) {$U_{-\mu}(n) \equiv U_\mu^\dagger(n-\hat\mu)$};
    \node (F) at (-4.5, -0.7) {$U_{\mu}(n)$};
\end{tikzpicture}
\capt{Schematic representation of the link variables $U_{\mu}(n)$ and $U_{-\mu}(n)$ on the lattice.}
\label{fig:links}
\end{figure}
 
With these results we can write a gauge invariant lattice fermion action as:
\beq
    S_F [\psi,\bar\psi] = a^4 \sum_{n\in\Lambda} \bpsi(n) \left[ \sum_{\mu=1}^4\gamma_\mu  \frac{U_\mu(n)\psi(n+\hat\mu) - U_{-\mu}(n)\psi(n-\hat\mu)}{2a} + m \psi(n) \right].
    \label{intro:lat_action}
\eeq

\subsection{The Gauge Transporter and the Wilson Loop}
For a more formal definition of the link variables we can look at the continuum gauge transporter $G(x,y)$, which is defined as the path-ordered product, denoted with $\mathcal{P}$, of a gauge field $A_\mu(x)$ along some curve $\mathcal{C}$ between two points $x$ and $y$ in spacetime:
\beq
    G(x,y) = \mathcal{P} \exp\left( ig_0\int_\mathcal{C} A_\mu(x')dx'^\mu  \right).
\eeq
An important thing to note now is that gauge transporter belongs to the gauge group, $\mathrm{SU}(3)$ for QCD, and not to the algebra, $\mathfrak{su}(3)$. 
Up to order $\mathcal{O}(a)$ the integral on the straight line connecting two points on the lattice can be approximated with $aA_\mu(n)$ for small lattice spacings. For convenience the gauge field is scaled by a factor $1/g_0$, giving: 
\beq
\tilde A_\mu(x) = \frac{1}{g_0} A_\mu(x),
\label{eq:fieldscale}
\eeq
but for notation's sake we'll drop the tilde from now on. One is then left with:
\begin{align}
    G(n,n+\hat\mu) &\equiv U_\mu(n) ~~= \exp(iaA_\mu(n)) = \mathds{1} + iaA_\mu(n) + \mathcal{O}(a^2) \label{eq:transporter}\\\nonumber
    G(n,n-\hat\mu) &\equiv U_{-\mu}(n) = \exp(-iaA_\mu(n)) = \mathds{1} - iaA_\mu(n-\hat\mu) + \mathcal{O}(a^2),
\end{align}
where in the second equation, for the negative direction, the approximation is taken at the arrival point instead of the initial one. With this definition is easy to see the relation we just stated earlier on the orientation change:
\beq
    U(n, n-\hat\mu) = U_{-\mu}(n) = U^\dagger_\mu(n).
\eeq
One can now define a Wilson loop as a gauge transporter that has the same start and final points. In the continuum case it can be shown, see for example ref. \cite{peskin}, that the trace of such an object is gauge invariant. On the lattice the minimal Wilson loop is just a square on a fixed plane $\mu\nu$, and we call this a Plaquette, which we label as $P_{\mu\nu}(n)$:
\begin{align}
    \label{plaquette}
P_{\mu\nu}(n) &= U_\mu(n) U_\nu(n+\hat\mu) U_{-\mu}(n+\hat\mu+\hat\nu) U_-{\nu}(n+\hat\nu)  \\\nonumber
              &= U_\mu(n) U_\nu(n+\hat\mu) U^\dagger_\mu(n+\hat\nu) U^\dagger_\nu(n).
\end{align}
A pictorial representation of this equation is given in \cref{fig:plaq}.
\begin{figure}[!htb]
    \centering
    \begin{tikzpicture}[node distance=1cm, remember picture] 

        \pgfmathsetmacro\xcoord{2.5};
        \pgfmathsetmacro\tiplen{0.4};
        % arrows
        \draw[->-=.55, line width=1.5pt, >=latex] (\xcoord,\xcoord) to (-\xcoord,\xcoord);
        \draw[->-=.55, line width=1.5pt, >=latex] (-\xcoord,\xcoord) to  (-\xcoord,-\xcoord);
        \draw[->-=.55, line width=1.5pt, >=latex] (-\xcoord,-\xcoord) to (\xcoord,-\xcoord);
        \draw[->-=.55, line width=1.5pt, >=latex] (\xcoord,-\xcoord) to  (\xcoord,\xcoord);

        \node (A) at (0, 3.1) {};%{$U_{-\mu}(n+\hat\mu+\hat\nu) = U^\dagger_{\mu}(n+\hat\nu)$};
        \node (B) at (-2.7, 0) {};%{$U_{-\nu}(n+\hat\mu) = U^\dagger_{\nu}(n+\hat\nu)$};
        \node (D) at (0, -3.1) {};%{$U_{\mu}(n)$};
        \node (C) at (2.7, 0) {};%{$U_{\nu}(n+\hat\mu) $};

        \draw (\xcoord, \xcoord) -- (\xcoord, \xcoord+\tiplen);
        \draw (\xcoord, \xcoord) -- (\xcoord+\tiplen, \xcoord);
        \draw (\xcoord, -\xcoord) -- (\xcoord, -\xcoord-\tiplen);
        \draw (\xcoord, -\xcoord) -- (\xcoord+\tiplen, -\xcoord);
        \draw (-\xcoord, -\xcoord) -- (-\xcoord, -\xcoord-\tiplen);
        \draw (-\xcoord, -\xcoord) -- (-\xcoord-\tiplen, -\xcoord);
        \draw (-\xcoord, \xcoord) -- (-\xcoord, \xcoord+\tiplen);
        \draw (-\xcoord, \xcoord) -- (-\xcoord-\tiplen, \xcoord);
        
        \filldraw (\xcoord, \xcoord) circle (0.05);
        \filldraw (\xcoord, -\xcoord) circle (0.05);
        \filldraw (-\xcoord, -\xcoord) circle (0.05);
        \filldraw (-\xcoord, \xcoord) circle (0.05); 
    \end{tikzpicture}

    \begin{tikzpicture}[node distance=1cm, remember picture, overlay] 
        \draw (A) node {$U_{-\mu}(n+\hat\mu+\hat\nu) = U^\dagger_{\mu}(n+\hat\nu)$};
        \draw [left](B)  node {$U_{-\nu}(n+\hat\mu) = U^\dagger_{\nu}(n+\hat\nu)$};
        \draw (D) node {$U_{\mu}(n)$};
        \draw [right](C) node {$U_{\nu}(n+\hat\mu) $};

        \draw[->, >=latex] (-5.7,1.1) to (-5.7,2.6);
        \draw[->, >=latex] (-5.7,1.1) to (-4.1,1.1);

        \node [left] (E) at (-5.7,2.6) {$x_\nu$};
        \node [below] (F) at  (-4.1,1.1) {$x_\mu$};
    \end{tikzpicture}
    \capt{The Plaquette, as defined on the lattice, is the oriented product of the link variables of a minimal square in the plane $\mu\nu$.}
    \label{fig:plaq}
\end{figure}

\subsection{The Wilson Gauge Action}
\label{sec:gluonAction}
In the continuum theory the Wilson loop is related to the square of the field strength tensor, the same term that enters the Lagrangian. The plaquette is therefore a good starting point for the creation of a discretized version of the Euclidean action for the gluon field.\\
Recalling the Baker-Campbell-Hausdorff (BCH) formula for the product of exponentials of matrices:
\beq
    \exp(A)\exp(B) = \exp\left( A+B+\frac{1}{2} [A,B]+\dots  \right),
\eeq
 and by inserting \cref{eq:transporter} repeatedly into the definition of the plaquette one has:
\begin{align}
    \nonumber P_{\mu\nu}(n) &=\exp\bigg[iaA_\mu(n) + iaA_\nu(n+\hat\mu) -iaA_\mu(n+\hat\nu) -iaA^\dagger_\nu(n) - \frac{a^2}{2}[A_\mu(n), A_\mu(n+\hat\mu)] \\
    & - \frac{a^2}{2}[A_\nu(n+\hat\nu), A_\nu(n)] + \frac{a^2}{2}[A_\mu(n), A_\nu(n)] + \frac{a^2}{2}[A_\nu(n+\hat\mu), A_\mu(n+\hat\nu)] \label{superlong} \\ 
    \nonumber & + \frac{a^2}{2}[A_\mu(n), A_\mu(n+\hat\nu)] + \frac{a^2}{2}[A_\nu(n+\hat\mu), A_\nu(n)] + \mathcal{O}(a^3)\bigg].
\end{align}
Now the terms that are shifted from the site $n$ can be expanded as:
\beq
    A_\mu(n+\hat\nu) = A_\mu(n) + a\partial_\nu A_\mu(n) +  \mathcal{O}(a^2)
\eeq
and with this substitution most terms cancel and one gets:
\begin{align}
    \label{plaq:expanded}
    P_{\mu\nu}(n) &=\exp\left[ ia^2(\partial_\mu A_\nu(n) - \partial_\nu A_\mu(n) + i[A_\mu(n),A_\nu(n)]) + \mathcal{O}(a^3)  \right]\\\nonumber
                  &=\exp\left[ ia^2G_{\mu\nu}(n) + \mathcal{O}(a^3)  \right],
\end{align}
where we recovered the gauge field strength tensor using $G_ {\mu\nu} =\partial_\mu A_\nu - \partial_\nu A_\mu + i[A_\mu,A_\nu]$. The relation found in \cref{plaq:expanded} can be used to build the euclidean lattice action term for the gluons. In particular, to be able to recover the continuum action, we would like a term of the form:
\beq
    S_G[U] = \frac{a^4}{2g^2}\sum_{n\in\Lambda}\sum_{\mu\nu} \Tr (G_{\mu\nu}(n)^2) \xrightarrow{a\rightarrow 0}  \frac{1}{4g^2} \int d^4x G_{\mu\nu}(x)^2 = S_G[A].
\eeq 
Note that the $1/g_0^2$ term, which could seem strange since it is not in the known continuum Lagrangian, is given by the scaling of the field in \cref{eq:fieldscale}. Up to order $\mathcal{O}(a^2) $ this can be obtained, in terms of the link variables, by:
\beq
    S_G[U] = \frac{2}{g^2}\sum_{n\in\Lambda}\sum_{\mu<\nu} \text{Re} \Tr (\mathds{1} - P_{\mu\nu}(n)).
    \label{wilsonaction}
\eeq 
Higher order corrections to this action can be computed analytically by considering higher orders in the BCH expansion in \cref{superlong} and in the exponential expansion in \cref{plaq:expanded}.

\subsection{Lattice Fermions} 
The discretization of fermions as discussed in \cref{lat:ferm_naive} is incomplete as it leads to unphysical results. It becomes evident if one considers the Fourier transform of the propagator. Let's first rewrite \cref{intro:lat_action} in a more compact way, introducing the lattice Dirac propagator $M_{xy}[U]$:
\beq
    S_F [\psi,\bar\psi] = a^4\sum_{n\in\Lambda} \bpsi(x)M_{xy}[U]\psi(y),
\eeq
with
\beq
    M_{xy}[U]= \sum_{\mu=1}^4\gamma_\mu  \frac{U_\mu(x)\delta_{x,(y-\hat\mu)} - U^\dagger_{-\mu}(x)\delta_{x,(y+\hat\mu)}}{2a} + m \delta_{x,y} .
\eeq
In momentum space the propagator becomes:
\beq
    \tilde M_{pq}= \delta(p-q)\tilde M(p)~~~~\text{where}~~~~\tilde M(p) = m\mathds{1} + \frac{i}{a}\sum_{\mu=1}^4\gamma_\mu\sin(p_\mu a).
\eeq
In order to calculate the inverse of the propagator in real space we need to invert the one in momentum space and perform an inverse Fourier transform. However, the inverse of the propagator in Fourier space has multiple poles, for example in the case of massless fermions:
\beq
\tilde M(p)^{-1}\bigg\rvert_{m=0} =  \frac{-ia\sum_{\mu}\gamma_\mu\sin(p_\mu a)}{\sum_{\mu}\sin^2(p_\mu a)}.
\label{inverse_prop}
\eeq
The problem vanishes for $a\rightarrow 0$, the continuum case, returning just one fermion type. On the contrary, for non-vanishing lattice spacings, multiple fermions, one for each pole, emerge from \cref{inverse_prop}. This is known as the ``doubling problem'' and the extra fermions are called doublers. The solution, proposed by Wilson \cite{wilson_confinement_1974}, is to modify the propagator adding a term that makes the poles in the inverse of the Fourier transformed propagator vanish. In momentum space the Wilson propagator is defined as:
\beq
\tilde M^W(p) = m\mathds{1} + \frac{i}{a}\sum_{\mu=1}^4\gamma_\mu\sin(p_\mu a) + \mathds{1}\frac{1}{a}\sum_{\mu=1}^4 (1-\cos(p_\mu a)).
\eeq
Now, by applying the inverse Fourier transform, one gets for coordinate space:
\beq
M^W_{xy}[U]= \frac{1}{2a}\sum_{\mu=\pm1}^{\pm4}(\mathds{1} -\gamma_\mu) U_\mu(x)\delta_{x,(y-\hat\mu)} + \left(m +\frac{4}{a}\right)\delta_{x,y} ,
\eeq
where the shorthand notation $\gamma_{-\mu} = -\gamma_\mu$ has been introduced. \\
The final form of the Wilson Fermion Action is:
\beq
    S_F [\psi,\bar\psi] = a^4\sum_{n\in\Lambda} \bpsi(x)M^W_{xy}[U]\psi(y).
    \label{intro:ferm_action}
\eeq

Now that all the needed information about the action and the fields is set, mainly through equations \cref{wilsonaction} and \cref{intro:ferm_action}, the picture of how to discretize QCD from the continuum Minkovskian spacetime to an euclidean spacetime lattice is complete.


\section{Path Integrals on the Lattice}
\label{sec:pathintegral}
To express expectation values and correlators on the lattice we employ the path integral formalism. The partition function, as we have seen earlier, is the path integral of the fields over the whole space of the exponential of the action. For the case of QCD the fields are $U$, $\psi$ and $\bpsi$:
\beq
	Z = \int \D\psi\D\bpsi\D U e^{-S_E[\psi,\bar{\psi},U] }  ,
\eeq
with the action being the sum of the gluonic and fermionic parts:
\beq
S_E[\psi,\bar{\psi},U]=S_G[U] + S_F[\psi,\bar{\psi}, U] = S_G[U] + \bpsi M[U] \psi.
\eeq
The immediate simplification is to integrate out the fermion fields. As in the continuum case one can perform the integration of the Grassmann-valued fields. \\
Grassmann numbers are defined to be anticommutative, meaning that for any two Grassmann numbers $a,b$ the relation $ab=-ba$ holds. A useful general property of these numbers regards the evaluation of multidimensional Gaussian integrals. For  set of Grassmann numbers $\theta_i$, their complex conjugates $\theta_i^*$ and a Hermitean matrix $K$ one can define the following integral:
\begin{align}
    \label{lattice:grassman}
    \int \D\Theta^*\D\Theta e^{-\Theta^* K \Theta } &= \left( \prod_i\int d\theta_i^*d\theta_i \right)  e^{-\theta_i^* K_{ij} \theta_j } =  \left( \prod_i\int d\theta_i^*d\theta_i \right)  e^{-\sum_i\theta_i^* k_i \theta_i } \\\nonumber
    &= \prod_i k_i = \det K .
\end{align} 
A more detailed proof of this relation can be found in \cite{peskin}. This result is very different from what one would get in the case of real numbers, where the analog integral is that of multi-variable gaussian:
\beq
    \prod_i^n\left(\int  dx_i^2\right) \left( \prod_{ij}^n x_iK_{ij}x_j  \right) = \frac{(2\pi)^n}{\det (K)}.
\eeq
With the result of \cref{lattice:grassman} we can simplify greatly the partition function:
\beq
	Z = \int \D\psi\D\bpsi\D U e^{-S[\psi,\bar{\psi},U] }  = \int \D U e^{-S_G[U] } \det M[U] .
\eeq
In a similar fashion as in statistical mechanics, the expectation value of an observable on the lattice can be computed, recalling the general form \cref{lattice:obs_base}, as:
\beq
    \langle O \rangle = \frac{1}{Z}  \int \D U ~O[\psi, \bpsi, U] e^{-S_G[U] } \det M[U] .
    \label{lattice:expectation}
\eeq
To compute fermionic correlators however, another relation of Grassmann integrals is needed:
\begin{align}
    \int \D\Theta^*\D\Theta \theta_a^*\theta_b e^{-\Theta^* K \Theta } &= \left( \prod_i\int d\theta_i^*d\theta_i \right) \theta_a^*\theta_b e^{-\theta_i^* K_{ij} \theta_j } =  \left( \prod_i\int d\theta_i^*d\theta_i \right) \theta_a^*\theta_b e^{-\sum_i\theta_i^* k_i \theta_i } \\\nonumber
    &=  (\det K ) (K^{-1})_{ab},
\end{align}
which allows us to write for the correlator of a fermion from a site $m$ to a site $n$ on the lattice:
\beq
    \langle O_n(\psi)O_m(\bpsi) \rangle = \frac{1}{Z}  \int \D U ~O_n[\psi, U]O_m[\bpsi, U] e^{-S_G[U] } (\det M[U]) (M[U]^{-1})_{mn}.
    \label{lattice:fermioncorrelator}
\eeq
This last result is crucial for the numerical evaluation of fermionic observables. For example, for a given ``source'' $\bpsi(x)$ and a ``sink'' $\psi(y)$ the correlator between the two can be computed via path integrals, but this requires inverting the fermion propagator matrix. \\
The above expressions for the partition function and the observables cannot be evaluated or simplified analytically any further, so the usual approach is to approximate the path integral numerically. The main idea is to create an ensemble of field configurations using the probability distribution function:
\beq
    P[U] = \frac{e^{-S_G[U]}\det M}{Z}.
\eeq 
Given an ensemble $\mathcal{U} = \{ U_1, U_2, \dots,U_N \}$ of configurations sampled with the PDF above, the integral in \cref{lattice:expectation} can be approximated as:
\beq
\langle O \rangle =  \int \D U~ P[U] O[\psi, \bpsi, U] \approx \frac{1}{N} \sum_{i=1}^N O(\psi, \bpsi, U_i), 
\eeq
so that the expectation value of $O$ is the average of the observable measure on the configurations of $\mathcal{U}$. The interesting, but difficult, problem that we have to face now is the choice of the ensemble. The solution is to use Monte Carlo methods, in particular to sample the space of the configurations using a Markov chain. In \cref{chap:code_design} a more detailed description of this problem, and a sketch of the implementation that has been developed to address it, will be presented.

\subsection{Pure Gauge Field Theory}
Computing full QCD on the lattice is computationally expensive, mainly due to the integration of fermions via the determinant. From a numerical point of view, the determinant needs to be computed at every step of the Markov chain that is used to evaluate the path-integral. This operation affects the time cost of sampling the configuration space dramatically. A first approach is to neglect the determinant completely, considering it constant. This, on a deeper level, effectively means removing dynamical fermions, freezing them to the lattice sites. This approximation is usually referred to as ``quenched QCD'', or QQCD. \\
The properties of this theory, that is then reduced to a simple Yang-Mills theory with zero fermion flavors, are still interesting to study and have historically played a very important role, being the only accessible simulation until sufficient computing power became available. For example, from \cref{beta:qcd} one can see that the theory presents confinement and asymptotic freedom. 

\subsection{Observables}
\label{sec:observables}
On the lattice, given the transformation \cref{link_transformation}, any product of link variables that starts and ends at the same site, a closed loop, is gauge invariant. The average values of these objects over the whole lattice can be linked to physical observables, for example the field tensor. In a more general form any observable $L[U]$ of the type:
\beq
    L[U] = \Tr \left[ \prod_{(n,\mu)\in\mathcal{P}} U_\mu(n)\right],
\eeq
where $\mathcal{P}$ is a closed path of links on the lattice, is a gauge invariant object.

\subsubsection{Plaquette}
The simplest observable, which we have already encountered upon defining the Wilson action in \cref{wilsonaction}, is the plaquette. This is the minimal closed loop on the lattice and its expectation value is related to the action. For each lattice site there are 6 possible plaquettes to be computed, given all the combinations of euclidean indices. One can consider the average value of the plaquette on a lattice field configuration as and observable related to the value of the action of the field, and consequently to the coupling constant. A proper definition of the observable, computed on a lattice $\Lambda$ is:
\beq
    P[U] = \frac{1}{6|\Lambda|}\sum_{n\in\Lambda}\sum_{\mu<\nu}P_{\mu\nu}(n),
\eeq
where $P_{\mu\nu}(n)$ is the one defined in \cref{plaquette} and $|\Lambda|$ is the notation that we introduce for the total number of lattice sites. \\
% The plaquette expectation value can be used as a tool to determine when the Markov Chain that is used to generate the ensemble, to measure  other observables, has reached a stationary point of the action. In particular, in \cref{sec:thermalization}, the use of the plaquette for checking thermalization of the chain is discussed.

\subsubsection{Energy Density}
The energy of the field is proportional to the square of the field tensor, so the energy density is in particular:
\beq
    E[U] = -\frac{1}{4|\Lambda|}\sum_{n\in\Lambda}\sum_{\mu<\nu}\text{Re}\Tr(G_{\mu\nu}(n)G_{\mu\nu}(n)).
    \label{eq:energy}
\eeq
In order to estimate this quantity, one has to compute the field tensor at every lattice site, square it and sum over the whole space. It is also usually normalized by the lattice volume (the number of sites) to get the density. The simplest definition of the field tensor is $G_{\mu\nu}^{(plaq)}(n) = \mathds{1} - P_{\mu\nu}(n)$ but this is not very accurate. A more symmetric definition can be obtained by the ``clover leaf'' field tensor, that is the average of all the plaquettes of a same plane $\mu\nu$ that start from a given lattice site taken all with the same orientation. To better understand the clover term, in figure \cref{fig:clover} a representation is given.

\begin{figure}[!htb]
    \centering
    \begin{tikzpicture}[node distance=1cm, remember picture] 

        \pgfmathsetmacro\len{2.5};
        \pgfmathsetmacro\start {0.2};

        % ++
        \draw[->-=.55, line width=1.5pt, >=latex] (\start,\start) to (\start+\len,\start);
        \draw[->-=.55, line width=1.5pt, >=latex] (\start+\len,\start) to  (\start+\len,\start+\len);
        \draw[->-=.55, line width=1.5pt, >=latex] (\start+\len,\start+\len) to (\start,\start+\len);
        \draw[->-=.55, line width=1.5pt, >=latex] (\start,\start+\len) to (\start, \start);

        %+-
        \draw[-<-=.55, line width=1.5pt, >=latex] (\start,-\start) to (\start+\len,-\start);
        \draw[-<-=.55, line width=1.5pt, >=latex] (\start+\len,-\start) to  (\start+\len,-\start-\len);
        \draw[-<-=.55, line width=1.5pt, >=latex] (\start+\len,-\start-\len) to (\start,-\start-\len);
        \draw[-<-=.55, line width=1.5pt, >=latex] (\start,-\start-\len) to (\start, -\start);

        %--
        \draw[-<-=.55, line width=1.5pt, >=latex] (-\start,-\start) to (-\start,-\start-\len);
        \draw[-<-=.55, line width=1.5pt, >=latex] (-\start,-\start-\len) to  (-\start-\len,-\start-\len);
        \draw[-<-=.55, line width=1.5pt, >=latex] (-\start-\len,-\start-\len) to (-\start-\len,-\start);
        \draw[-<-=.55, line width=1.5pt, >=latex] (-\start-\len,-\start) to (-\start, -\start);

        %-+
        \draw[->-=.55, line width=1.5pt, >=latex] (-\start,\start) to (-\start,\start+\len);
        \draw[->-=.55, line width=1.5pt, >=latex] (-\start,\start+\len) to  (-\start-\len,\start+\len);
        \draw[->-=.55, line width=1.5pt, >=latex] (-\start-\len,\start+\len) to (-\start-\len,\start);
        \draw[->-=.55, line width=1.5pt, >=latex] (-\start-\len,\start) to (-\start, \start);

        \filldraw (0,0) circle (0.1);
    \end{tikzpicture}


    \begin{tikzpicture}[node distance=1cm, remember picture, overlay] 
        \pgfmathsetmacro\lenarrow{1.2};
        \pgfmathsetmacro\startarrowx{-5.7};
        \pgfmathsetmacro\startarrowy{0.5};
        \draw[->, >=latex] (\startarrowx, \startarrowy) to (\startarrowx, \startarrowy+\lenarrow);
        \draw[->, >=latex] (\startarrowx, \startarrowy) to  (\startarrowx+\lenarrow, \startarrowy);

        \node [left] (E) at  (\startarrowx, \startarrowy+\lenarrow) {$x_\nu$};
        \node [below] (F) at (\startarrowx+\lenarrow, \startarrowy) {$x_\mu$};
    \end{tikzpicture}
    \capt{Schematic representation of the symmetric definition of the clover term $G_{\mu\nu}^{(clover)}(n)$ on the lattice. The field strength tensor for a plane $\mu\nu$ at a lattice site is given by the average of all the for plaquettes that start from a site $n$, all taken with the same orientation.}
    \label{fig:clover}
\end{figure}

In terms of link variables this is equal to:
\begin{align}
    G_{\mu\nu}^{(clover)}(n)=\frac{1}{4}\bigg[  &U_\mu(n)U_\nu(n+\hat\mu)U^\dagger_\mu(n+\hat\nu)U^\dagger_\nu(n)\\\nonumber
    + &U_\nu(n)U^\dagger_\mu(n+\hat\nu)U^\dagger_\nu(n-\hat\mu)U_\mu(n-\hat\mu)\\\nonumber
    + &U^\dagger_\mu(n-\hat\mu)U^\dagger_\nu(n-\hat\mu-\hat\nu)U_\mu(n-\hat\mu-\hat\nu)U_\nu(n-\hat\nu)\\\nonumber
    + &U^\dagger_\nu(n-\hat\nu)U_\mu(n-\hat\nu)U_\nu(n+\hat\mu-\hat\nu)U^\dagger_\mu(n) \bigg].
\end{align}
 

\subsubsection{Topological Charge}
The gauge field in QCD exhibits particular topological properties that are believed to have important physical implications \cite{witten_current_1979,di_giacomo_topology_1997}. The topological charge is an integer quantum number of the field in the continuum, similar to the concept of winding-number. In general fields with different topological charge are separated, meaning that no unitary transformation can change the topological properties of the field. However, there exist non-zero instanton modes \cite{schafer_instantons_1998} that allow tunneling between different ``topological sectors".
On the lattice certain definitions can be used to reproduce the continuum properties, especially for the so called ``topological susceptibility'', that is the second moment of the distribution of the topological charge, which seems to be independent from the definitions of the base observable \cite{alexandrou_comparison_2017,gross_qcd_1981}. 

% \NOTE{CHECK...}
% In the continuum topological sectors, regions of space with same charge, are separated from each other, on the lattice through discretization effects the behavior is different, with instantons that allow tunneling between sectors \cite{gross_qcd_1981}.\\ 


The topological charge is the integral over all spacetime of the topological charge density
\beq
    Q = \int d^4xq(x),
\eeq
where
\beq 
    q(x)=\frac{1}{32\pi^2}\Tr(G_{\mu\nu}(x)\tilde G^{\mu\nu}(x)) .
\eeq 
Here with $\tilde G^{\mu\nu}$ the dual of the gauge field is intended and its definition is:
\beq
\tilde G^{\mu\nu} (x)= \frac{1}{2}\epsilon^{\mu\nu\rho\sigma} G^{\rho\sigma}(x),
\eeq
with $\epsilon^{\mu\nu\rho\sigma}$ being the anti-symmetric Levi-Civita symbol. We can now see the link between this observable and the $\theta$-term onf the QCD Lgrangian we mentioned in \cref{intro:symmetry}.\\
The topological charge can be estimated on the lattice, in the simplest way, by using the same clover definition of the field strength tensor we used before for the energy density:
\beq
    Q[U]=\frac{1}{64\pi^2} \sum_{n\in\Lambda}\sum_{\mu<\nu}\epsilon^{\mu\nu\rho\sigma}\Tr[G_{\mu\nu}^{(clover)}(n)G_{\rho\sigma}^{(clover)}(n)].
    \label{eq:topc}
\eeq
As the lattice spacing is reduced, approaching the continuum, the topological sectors get more and more separated, preventing tunneling between them. This makes the Markov chain used to generate the ensemble less efficient in terms of growing autocorrelation times of the observables, as will be shown in \cref{sec:testautocorr}.\\
A derived quantity of great interest is the topological susceptibility $\chi$, that is proportional to the expectation value of the square of the topological charge:
\beq
    \chi = \frac{1}{a^4|\Lambda|}\langle Q^2 \rangle.
    \label{eq:topsus}
\eeq
This observable is particularly important for the properties of instantons on the lattice and its value can be related via the Witten-Veneziano formula \cite{witten_current_1979} to the mass of the $\eta'$ meson.

\section{Modern Lattice QCD Calculations}
Lattice QCD, after in 40 years of time made huge improvements. However, still today some fascinating challenges are open for investigations in the field. \\
The state of the art calculations in Lattice QCD involve improved fermion and gluon actions. For the gluon sector the improvements are based on finding linear combinations of gauge invariant Wilson loops, such as $2\times 1$, $2\times 2$ and $3\times 1$ rectangles to systematically remove the higher order in the right-hand side of \cref{plaq:expanded}. For fermions the improvements consist on actions that preserve certain symmetries more than others, like chiral fermion actions \cite{degrand_chiral_2007}.\\
On the algorithmic side, the currently most used method to generate gauge field configuration is the Hybrid Monte Carlo method \cite{alles_hybrid_1996}, which combines stochastic sampling with hamiltonian dynamic updates in the  gauge field space. Other interesting algorithmic problems regard the calculation of fermion determinants, which typically involve very large sparse matrix diagonalization.\\
A fundamental concept in Lattice calculations is to recover the continuum limit in a controlled manner at the end of the analysis. The error sources that come into play when discretizing spacetime on a fixed lattice are:
\begin{itemize}
    \item Finite Volume Effects: caused by the non infinite domain of the simulation. Usually periodic boundary conditions are implemented to simulate infinite space, but one should always check if the total lattice volume is large enough, especially when dealing with large systems such as multiple baryons.
    \item State Isolation and Signal to Noise: when computing correlators between two points on the lattice it is possible to extract energy  states of hadrons by looking at the exponential decay of the correlator in euclidean time. However this problem is largely affect by signal to noise problems making it hard to extract excited states.
    \item Chiral Limit: by which is intended the limit for which the masses of the quarks, and consequently of the computed hadrons, on the lattice approach the physical masses \cite{fukaya_chiral_2008}. One might question why this is even the case and that calculations should be performed at the physical masses only. However, the fermion propagator matrix becomes harder to invert, because of vanishing eigenvalues, the lower he mass of the quark is. This makes the numerical algorithms that should diagonalize it slower to converge. The usual approach is to perform calculations on a set of pion masses (this is the usual reference for this problem) and afterwards the limit for $m_\pi^{(lat)} \rightarrow m_\pi^{(phys)}$ is taken to reproduce the physical observables.
    \item Continuum Limit: perhaps the most obvious thing to do to improve Lattice QCD calculations is to take finer and finer lattice spacings, to approach the continuum case. It should be noted however that this procedure has one major downside. As the lattice approaches the continuum case the autocorrelation of observables in a given ensemble increases rapidly. This behavior, known as ``critical slow-down'' \cite{schaefer_critical_2011} depends on the observable itself, some are more affected then others  The integrated autocorrelation time for the topological charge, for example, is believed to have either  power-law with a large exponent or even an exponential relation to the lattice spacing. 
\end{itemize}
As one can notice, Lattice QCD is still an open field of research in many aspects: the algorithmic/numerical, the extraction and improvement of the calculations of physical quantities and the theoretical model itself. 

